---
title: "Untitled"
author: "Nicolas May"
date: "2/26/2019"
output: 
  html_document:
    theme: sandstone
    highlight: zenburn
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Requirements

Load useful packages.
```{r}
library(dplyr)
library(stringr)
library(readr)
library(FuzzyDateJoin)
```

Load useful globals and functions.
```{r}
source("~/Desktop/config.R")
source("~/Desktop/helpers.R")
```


# Data

## Get Data

### Get UMMAP IDs of interest
```{r}
reqd_ids <- readxl::read_excel("MADC_data_query.xlsx") %>% 
  # keep only distinct UM IDs
  distinct(`ID Ext`) %>% 
  # rename the field
  rename(ptid = `ID Ext`) %>% 
  # extract the four relevant digits of the UM ID
  mutate(ptid = str_replace(ptid, "^hlp17umm0", "")) %>%
  mutate(ptid = str_replace(ptid, "_\\d{5}$", "")) %>% 
  # prepend "UM0000" to four relevant digits of UM ID
  mutate(ptid = str_glue("UM0000{ptid}", ptid = ptid)) %>% 
  # keep only IDs with pattern `UMxxxxxxxx`
  filter(str_detect(ptid, "^UM\\d{8}$")) %>%
  # df => vector
  pull() %>% 
  # coerce to simple character vector
  as.character() %>% 
  # sort 
  sort()
```

### Get NIH Toolbox data

```{r}
df_tb <- 
  read_csv(
    file = "./TB_renormed.csv",
    col_types = cols(
      .default        = col_double()
      , subject_id    = col_character()
      , PIN           = col_character()
      , tb_admin_date = col_datetime(format = "%m/%d/%y")
      , Language      = col_character()
      , Age           = col_integer()
      , Gender        = col_integer()
      , Race          = col_integer()
      , Ethnic        = col_integer()
      , Education     = col_integer()
    )) %>% 
  rename(
    ptid      = subject_id
    , tb_date = tb_admin_date
  ) %>% 
  select(
    -Language
    , -Age
    , -Gender
    , -Race
    , -Ethnic
    , -Education
  )

# Remove columns without any data
df_tb_keep_cols <- purrr::map_lgl(df_tb, ~ any(!is.na(.x)) )
df_tb <- df_tb[, df_tb_keep_cols]
```

### Get CogState data

```{r}
df_cs <- 
  read_csv(
    file = "./cogstate_complete.csv",
    col_types = cols(
      .default     = col_double()
      , subject_id = col_character()
      , BDate      = col_date(format = "%Y-%m-%d")
      , BYear      = col_integer()
      , Hand       = col_character()
      , Sex        = col_character()
      , Educ       = col_character()
      , TDate      = col_date(format = "%Y-%m-%d")
      , TTime      = col_datetime()
    )) %>% 
  rename(
    ptid      = subject_id
    , cs_date = TDate
  ) %>% 
  select(
    -BDate
    , -BYear
    , -Hand
    , -Sex
    , -Educ
    , -TTime
  )

# Remove columns without any data
df_cs_keep_cols <- purrr::map_lgl(df_cs, ~ any(!is.na(.x)) )
df_cs <- df_cs[, df_cs_keep_cols]
```

### Define relevant data fields

#### UDS 3

* Age (DOB)
* Sex
* UDS Dx
* MoCA (total)

```{r}
# Form Header (IVP)
fields_u3_hd_raw <-
  c(
    "ptid"
    , "form_date"
    , "dob"       # for deriving age
  )
# Form A1 (IVP, FVP, TVP)
fields_u3_a1_raw <-
  c(
    "sex"
  ) %>% c(., paste0("fu_", .), paste0("tele_", .))

# Form C2 (IVP, FVP)
fields_u3_c2_raw <-
  c(
    "mocatots"
  ) %>% c(., paste0("fu_", .))

# Form D1 (IVP, FVP, TVP)
fields_u3_d1_raw <-
  c(
    "normcog"    # NL
    , "demented" # Demented
    , "amndem"
    , "pca"
    , "ppasyn"
    , "ftdsyn"
    , "lbdsyn"
    , "namndem"
    , "mciamem"  # MCI
    , "mciaplus" # MCI
    , "mcinon1"  # MCI
    , "mcinon2"  # MCI
    , "impnomci" # Impaired not MCI
    , "alzdis"
    , "alzdisif" 
    , "lbdis"
    , "lbdif" 
    , "msa"
    , "msaif"
    , "psp"
    , "pspif"
    , "cort"
    , "cortif"
    , "ftldmo" 
    , "ftldmoif"
    , "ftldnos"
    , "ftldnoif"
    , "cvd" 
    , "cvdif" 
    , "esstrem"
    , "esstreif"
    , "downs" 
    , "downsif"
    , "hunt"
    , "huntif"
    , "prion"
    , "prionif"
    , "brninj"
    , "brninjif"
    , "hyceph"
    , "hycephif"
    , "epilep" 
    , "epilepif"
    , "neop"
    , "neopif" 
    , "hiv"
    , "hivif"
    , "dep"
    , "depif" 
    , "bipoldx"
    , "bipoldif"
    , "schizop"
    , "schizoif"
    , "anxiet"
    , "anxietif"
    , "delir"
    , "delirif"
    , "ptsddx"
    , "ptsddxif"
    , "othpsy"
    , "othpsyif"
    , "alcdem"
    , "alcdemif"
    , "impsub"
    , "impsubif"
    , "dysill"
    , "dysillif"
    , "meds"
    , "medsif"
  ) %>% c(., paste0("fu_", .), paste0("tele_", .))

# Combine and collapse fields
fields_u3_raw <-
  c(
    fields_u3_hd_raw
    , fields_u3_a1_raw
    , fields_u3_c2_raw
    , fields_u3_d1_raw
  )

fields_u3 <- fields_u3_raw %>% paste(collapse = ",")

# Forms Toolbox & CogState
forms_u3_raw <-
  c(
    # "tool_box"
    # , "cogstate"
    "ipad_tool_box"
  )

# Combine and collapse forms
forms_u3 <- forms_u3_raw %>% paste(collapse = ",")
```

### Retrieve data via REDCAP API

#### UDS 3

```{r}
json_u3 <- rc_api_get(uri    = REDCAP_API_URI,
                      token  = REDCAP_API_TOKEN_UDS3n,
                      fields = fields_u3,
                      forms  = forms_u3)
```

```{r}
df_u3 <- jsonlite::fromJSON(json_u3) %>% na_if("")
```

## Process Data

### Clean data

#### UDS 3

```{r}
df_u3_cln <- df_u3 %>% 
  # deselect useless field(s)
  select(-redcap_event_name) %>% 
  # remove DDE records (--1, --2)
  filter(str_detect(ptid, "^UM\\d{8}$")) %>% 
  # remove missing `form_date`s
  filter(!is.na(form_date)) %>% 
  # keep only UM IDs identified by data requestor
  filter(ptid %in% reqd_ids)
```

#### NIH Toolbox

```{r}
df_tb_cln <- df_tb %>% 
  # Deselect useless field(s)
  select(-PIN) %>% 
  # clean out missing UM IDs
  filter(!is.na(ptid)) %>% 
  # clean out missing `tb_date`s
  filter(!is.na(tb_date)) %>% 
  # keep only UMMAP IDs
  filter(str_detect(ptid, "^UM\\d{8}$"))
```

#### CogState

```{r}
df_cs_cln <- df_cs %>% 
  # clean out missing UM IDs
  filter(!is.na(ptid)) %>% 
  # clean out missing `cs_date`s
  filter(!is.na(cs_date)) %>% 
  # keep only UMMAP IDs
  filter(str_detect(ptid, "^UM\\d{8}$"))
```

### Mutate data

#### UDS 3

```{r}
df_u3_cln_mut <- df_u3_cln %>% 
  # coerce `form_date` to date
  mutate(form_date = as.Date(form_date)) %>% 
  # coalesce IVP/FVP/TVP fields
  coalesce_ift_cols() %>% 
  # simplify UDS diagnosis fields
  mutate(uds_dx_der = case_when(
    normcog  == 1 ~ "NL",
    demented == 1 & amndem == 1 ~ 
      "Amnestic multidomain dementia syndrome",
    demented == 1 & pca == 1 ~
      "Posterior cortical atrophy syndrome",
    demented == 1 & ppasyn == 1 ~
      "Primary progressive aphasia (PPA) syndrome",
    demented == 1 & ftdsyn == 1 ~
      "Behavioral variant FTD (bvFTD) syndrome",
    demented == 1 & lbdsyn == 1 ~
      "Lewy body dementia syndrome",
    demented == 1 & namndem == 1 ~
      "Non-amnestic multidomain dementia, not PCA, PPA, bvFTD, or DLB syndrome",
    demented == 0 & mciamem  == 1 ~ "MCI",
    demented == 0 & mciaplus == 1 ~ "MCI",
    demented == 0 & mcinon1  == 1 ~ "MCI",
    demented == 0 & mcinon2  == 1 ~ "MCI",
    demented == 0 & impnomci == 1 ~ "Impaired no MCI",
    TRUE ~ NA_character_
  )) %>% 
  # simplify UDS etiology fields
  mutate(uds_prim_etio = case_when(
    alzdis   == 1 & alzdisif == 1 ~ "AD",
    lbdis    == 1 & lbdif    == 1 ~ "LBD",
    msa      == 1 & msaif    == 1 ~ "MSA",
    psp      == 1 & pspif    == 1 ~ "PSP",
    cort     == 1 & cortif   == 1 ~ "CBD",
    ftldmo   == 1 & ftldmoif == 1 ~ "FTLD with motor neuron disease",
    ftldnos  == 1 & ftldnoif == 1 ~ "FTLD NOS",
    cvd      == 1 & cvdif    == 1 ~ "Vascular brain injury",
    esstrem  == 1 & esstreif == 1 ~ "Essential tremor",
    downs    == 1 & downsif  == 1 ~ "Down syndrome",
    hunt     == 1 & huntif   == 1 ~ "Huntington's disease",
    prion    == 1 & prionif  == 1 ~ "Prion disease",
    brninj   == 1 & brninjif == 1 ~ "Traumatic injury",
    hyceph   == 1 & hycephif == 1 ~ "Normal-pressure hydrocephalus",
    epilep   == 1 & epilepif == 1 ~ "Epilepsy",
    neop     == 1 & neopif   == 1 ~ "CNS neoplasm",
    hiv      == 1 & hivif    == 1 ~ "HIV",
    TRUE ~ NA_character_
  )) %>% 
  # simplify UDS conidtion fields
  mutate(uds_condition = case_when(
    dep      == 1 & depif    == 1 ~ "Active depression",
    bipoldx  == 1 & bipoldif == 1 ~ "Bipoloar disorder",
    schizop  == 1 & schizoif == 1 ~ "Schizophrenia",
    anxiet   == 1 & anxietif == 1 ~ "Anxiety disorder",
    delir    == 1 & delirif  == 1 ~ "Delirium",
    ptsddx   == 1 & ptsddxif == 1 ~ "PTSD",
    othpsy   == 1 & othpsyif == 1 ~ "Other psychiatric disease",
    alcdem   == 1 & alcdemif == 1 ~ 
      "Cognitive impairment due to alcohol abuse",
    impsub   == 1 & impsubif == 1 ~ 
      "Cognitive impairment due to other substance abuse",
    dysill   == 1 & dysillif == 1 ~
      "Cognitive impairment due to systemic disease/medical illness",
    meds     == 1 & medsif   == 1 ~
      "Cognitive impairment due to medications",
    TRUE ~ NA_character_
  )) %>% 
  # deselect useless columns
  select(
    -normcog,
    -mciamem, -mciaplus, -mcinon1, -mcinon2,
    -impnomci,
    -alzdis, -alzdisif,
    -lbdis, -lbdif,
    -msa, -msaif,
    -psp, -pspif,
    -ftldmo, -ftldmoif, -ftldnos, -ftldnoif,
    -demented,
    -amndem, -pca, -ppasyn, -ftdsyn, -lbdsyn, 
    -cvd, -cvdif, -brninj, -brninjif,
    -dep, -depif, -anxiet, -anxietif, -othpsy, -othpsyif, 
    -impsub, -impsubif, -dysill, -dysillif, -meds, -medsif
  ) %>% 
  # calculate age
  calculate_age(dob, form_date) %>% 
  select(-dob, -age_years, -age_units) %>% 
  # reorder columns
  select(ptid, form_date, sex, age_exact, 
         uds_dx_der, uds_prim_etio, uds_condition,
         everything())
```

#### NIH Toolbox

```{r}
df_tb_cln_mut <- df_tb_cln %>% 
  mutate(tb_date = as.Date(tb_date))
```

#### CogState

```{r}
df_cs_cln_mut <- df_cs_cln %>% 
  mutate(cs_date = as.Date(cs_date))
```

### Join Data

Left join: df_u3_cln_mut, df_tb_cln, df_cs_cln

```{r}
df_u3_tb_cs <-
  outer_left(x = df_u3_cln_mut,        y = df_tb_cln_mut,
             x_id_col = "ptid",        y_id_col = "ptid",
             x_date_col = "form_date", y_date_col = "tb_date",
             x_intvl_less = 120L, x_intvl_more = 120L,
             keep_y_id = FALSE) %>%
  rename(ptid = ptid_x) %>% 
  outer_left(x = .,                    y = df_cs_cln_mut,
             x_id_col = "ptid",        y_id_col = "ptid",
             x_date_col = "form_date", y_date_col = "cs_date",
             x_intvl_less = 30L * 6L, x_intvl_more = 30L * 6L, # +/- 6 months
             keep_y_id = FALSE) %>% 
  rename(ptid = ptid_x)
```

Clean out the columns that have no data.

```{r}
df_u3_tb_cs <- df_u3_tb_cs %>% 
  select_if(
    function(x) {
      any(!is.na(x))
    })
```


## Write Data

```{r}
write_csv(df_u3_tb_cs, "df_u3_tb_cs.csv", na = "")
```


# Data Dictionaries

## Get Data Dictionaries

Get field names.

```{r}
names_df_u3_tb_cs <- names(df_u3_tb_cs) %>% tolower()

# names_df_u3_tb_cs[str_detect(names_df_u3_tb_cs, "_log$")]
# names_df_u3_tb_cs[str_detect(names_df_u3_tb_cs, "_raw$")]

# names_df_u3_tb_cs <- names_df_u3_tb_cs %>% 
#   str_replace("_log$", "log") %>% 
#   str_replace("_raw$", "raw")

# fields_dd <- names_df_u3_tb_cs %>% paste(collapse = ",")
# using the `fields` argument in the API call doesn't work for some reason ???
```

Get data dictionary via REDCap API.

```{r}
json_dd_u3_tb_cs <- RCurl::postForm(
  uri          = REDCAP_API_URI,
  token        = REDCAP_API_TOKEN_UDS3n,
  content      = 'metadata',
  format       = 'json',
  returnFormat = 'json'
)

df_dd_u3_tb_cs <- jsonlite::fromJSON(json_dd_u3_tb_cs)
```

## Process Data Dictionaries

Keep only relevant fields

```{r}
df_dd_u3_tb_cs_cln <- df_dd_u3_tb_cs %>% 
  filter(field_name %in% names_df_u3_tb_cs)
```

## Write Data Dictionaries

```{r}
readr::write_csv(df_dd_u3_tb_cs_cln, "DataDictionary_UDS3.csv", na = "")
```


```{r echo=FALSE}
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
